Learning constraints in spreadsheets and tabular data

### 2016-05-27 Paper/ECAI/tabular.tex
Fixes

A block has the following properties:
\begin{itemize}
\item \textit{type}: a block is type-consistent, so it has a type
\item \textit{table}: the table that the block belongs to
\item \textit{orientation}: either row-oriented or column-oriented
\item \textit{size}: the number of vectors a block contains
\item \textit{length}: the length of its vectors; as all vectors are from the same table, they always have the same length.
\\\\
\luc{these are the vectors along \dots along}
\item \textit{rows}: this is orientation specific, for row-based blocks this is the size, for column-based blocks the length
\item \textit{columns}: also orientation specific, for row-based blocks this is the length, for column-based blocks the size
\end{itemize}

@@

  \item \CSignature: $\sg_r$ has to be a single vector ($\textit{size}=1$) while $\mathbf{\sg_x}$ can be a block ($\textit{size}>=1$), which can be derived from the use of a normal or \textbf{bold} font. The two blocks have to be numeric. This constraint is orientation-specific, so it requires that the number of rows in $\mathbf{\sg_x}$ equals the length of $\sg_r$. While not strictly needed, we also add to the bias of this template that the number of columns to sum over is larger than $2$ as a block with two columns will also be captured by the $\eccalc{\sg_r}{\sg_1 + \sg_2}$ constraint;

  \item \CFunction: each value in the vector $\sg_{r}$ is obtained by summing over the corresponding row in $\mathbf{\sg_x}$. \luc{example where it holds}
\end{itemize}
\end{example}

@@

\subsubsection{Redundancies}
\label{sec:form:redundancies}
Depending on the application, some constraints in the solution to the tabular constraint learning problem may be considered \textit{redundant}. This is because constraints may be logically equivalent or may be implied by other constraints.
\luc{if you have permutation, alldiff is redundant.}

@@

Our proposed methodology contains the following steps:
\begin{enumerate}
\item Extract headerless tables from tabular data
\item Partition the tables into maximally contiguous, type-consistent \textit{superblocks}
\item Generate for each constraint template all valid argument assignments in two steps:
\begin{enumerate}
\item For each constraint $c$, generate all atoms $c(B_1, \ldots ,B_n)$ where each $B_i$ is a maximal superblock and the $B_i$ are \textit{compatible} (defined below) with the signature. 
\item For each generated atom $c(B_1, \ldots ,B_n)$ in 3a), find all valid $c(B'_1, \ldots, B'_n)$ such that for all $i$ holds $B'_i \sqsubseteq B_i$ and the signature and definition of $c(B'_1, \ldots, B'_n)$ are satisfied.
\end{enumerate}
\end{enumerate}
\luc{introduce the notion of a maximal superblock earlier ???}

@@

In step 3b, we start from an atom $c(B_1, \ldots, B_n)$ with maximal superblocks $B_i$, and generate and test all possible (sub)block assignments to arguments using the properties and actual data of the blocks. As we only have to consider the blocks $B'_i \sqsubseteq B_i$ contained in each superblock, this is typically much easier. For the above example, each of the vectors of $B_6$ will be considered as candidate for the left-hand side, and one can enumerate all subblocks for the right-hand side and verify the signature (e.g. at least size $4$) and test whether for each row the definition is satisfied. %In practice, this can be optimized further.
\\\\
\samuel{Section references not correct}
We now describe in more detail how the headerless tables are extracted and how the maximal superblocks are generated from them (step 1 and 2, Section 4.1), how the candidate superassignments are generated (step 3a, Section 4.2) and how the actual assignments are extracted from that (step 3b, Section 4.3). We then describe a number of optimizations that we perform to speed up steps 3 and 4 (Section 4.4).

@@

    \Procedure{LearnConstraints}{$\blocks$, $\temps$}
      \State $C \gets \emptyset$ %\Comment{The set of constraints}
      \ForAll{$s~\mathbf{in}~\temps$}
        %\State $n \gets$ number of arguments of template $s$
        \State $A \gets \generategroups(s, \blocks)$
        \ForAll{$(B_1, \dots, B_n) \in A$}
        \State $A' \gets \findassignment(s, (B_1,\dots,B_n))$
        \ForAll{$(B_1',\dots,B_n') \in A'$}
            \Comment \luc{if cs(B; …( satisfied/?}
            \State $C \gets C \cup \{ c_s(B_1', \dots, B_n') \}$
          \EndFor
        \EndFor
      \EndFor
      \State \Return $C$
    \EndProcedure
\end{algorithmic}
\caption{Learn tabular constraints}
\label{algo:tcl}
\end{algorithm}

\subsubsection{Generating superblock assignments (Step 3a)}
% Let there be $b$ blocks and let~$s$ have~$n$ arguments, then a simple generate-and-test method would have to test $b^n$ combinations (given that blocks can be repeated), which can grow very rapidly with $b$, depending on $n$.

@@

Hence, a simple generate-and-test will usually suffice.
%However, the actual implementation is specified for every constraint template separately, allowing different methods to be used based on the type of template.
%For this reason, the \sname system uses generate-and-test to implement $\findassignment$ for most constraint templates.
\luc{test = definition, why use another term ?}
Given a superblock assignment, all disjoint subassignments are generated taking into account the required size, columns or rows.
The disjoint subassignments are then sequentially tested using a template-specific procedure \textit{Test} and valid subassignments are returned.
The \textit{Test} procedure verifies the exact signature (e.g. subtypes will not have been checked yet) and checks whether the definition is satisfied for all data in the blocks.

\begin{algorithm}[tbh]
  \begin{algorithmic}
    \footnotesize
    \Procedure{Subassignments}{$s$, $(B_1, \dots, B_n)$}
    \State $n \gets$ number of arguments of template $s$
    \State $A_{sub} \gets \emptyset$
    \ForAll{$(B_1', \dots, B_n') \textbf{ where } \forall i: B_i' \sqsubseteq B_i  \land \textit{disjoint}_{j \neq i}(B_i', B_j')$}
      \If{$Test((B_1',\dots,B_n'))$}
      \Comment \luc{Test = def is satisfied}
        \State $A_{sub} \gets A_{sub} \cup \{ (B_1', \dots, B_n') \}$
      \EndIf
    \EndFor
\Return $A_{sub}$
\EndProcedure
\end{algorithmic}
\caption{Generate-and-test for $\findassignment$}
\label{algo:gen-test-subblocks}
\end{algorithm}

>>>>

A block has the following properties:
\begin{itemize}
\item \textit{type}: a block is type-consistent, so it has a type
\item \textit{table}: the table that the block belongs to
\item \textit{orientation}: either row-oriented or column-oriented
\item \textit{size}: the number of vectors a block contains
\item \textit{length}: the length of its vectors; as all vectors are from the same table, they always have the same length.
\item \textit{rows}: the number of rows in the block
(in row-oriented blocks this is equivalent to the size)
\item \textit{columns}: the number of columns in the block (in row-oriented blocks this is equivalent to the length)
\end{itemize}

@@

  \item \CSignature: $\sg_r$ has to be a single vector ($\textit{size}=1$) while $\mathbf{\sg_x}$ can be a block ($\textit{size}>=1$), which can be derived from the use of a normal or \textbf{bold} font. The two blocks have to be numeric. This constraint is orientation-specific, so it requires that the number of rows in $\mathbf{\sg_x}$ equals the length of $\sg_r$. While not strictly needed, we also add to the bias of this template that the number of columns to sum over is larger than $2$ as a block with two columns will also be captured by the $\eccalc{\sg_r}{\sg_1 + \sg_2}$ constraint; \samuel{check}

  \item \CFunction: each value in the vector $\sg_{r}$ is obtained by summing over the corresponding row in $\mathbf{\sg_x}$. 
  For example, $B_r = [12, 7]$ and $\mathbf{\sg_x}$ consists of columns: $[5, 15]$, $[7, -8]$.
\end{itemize}
\end{example}

@@

\subsubsection{Redundancies}
\label{sec:form:redundancies}
Depending on the application, some constraints in the solution to the tabular constraint learning problem may be considered \textit{redundant}. This is because constraints may be logically equivalent or may be implied by other constraints.

\luc{if you have permutation, alldiff is redundant.}

@@

Our proposed methodology contains the following steps:
\begin{enumerate}
\item Extract headerless tables from tabular data
\item Partition the tables into contiguous, type-consistent \textit{superblocks}
\item Generate for each constraint template all valid argument assignments in two steps:
\begin{enumerate}
\item For each constraint template~$s$, generate all assignments $(B_1, \ldots ,B_n)$ where each $B_i$ is a superblock and the $B_i$ are \textit{compatible} (defined below) with the signature of~$s$. 
\item For each generated assignment $(B_1, \ldots ,B_n)$ in 3a), find all valid constraints $s(B'_1, \ldots, B'_n)$ such that for all $i$ holds $B'_i \sqsubseteq B_i$ and the signature and definition of~$s$ are satisfied.
\end{enumerate}
\end{enumerate}
\luc{introduce the notion of a maximal superblock earlier ???}

@@

In step 3b, we start from an assignment $(B_1, \ldots, B_n)$ with superblocks $B_i$, and generate and test all possible (sub)block assignments to arguments using the properties and actual data of the blocks. As we only have to consider the blocks $B'_i \sqsubseteq B_i$ contained in each superblock, this is typically much easier. For the above example, each of the vectors of $B_6$ will be considered as candidate for the left-hand side, and one can enumerate all subblocks for the right-hand side and verify the signature (e.g. at least size $4$) and test whether for each row the definition is satisfied. %In practice, this can be optimized further.

We now describe in more detail how headerless tables are extracted (step 1, Section~\ref{sec:table_extraction}) and how superblocks are generated from them (step 2, Section~\ref{sec:make_groups}), how the candidate superblock assignments are generated (step 3a, Section~\ref{sec:algo:super}) and how the actual assignments are extracted from that (step 3b, Section~\ref{sec:algo:subgr}).
In Section~\ref{sec:opts} we describe a number of optimizations designed to speed up steps the algorithm.

@@

    \Procedure{LearnConstraints}{$\blocks$, $\temps$}
      \State $C \gets \emptyset$ %\Comment{The set of constraints}
      \ForAll{$s~\mathbf{in}~\temps$}
        %\State $n \gets$ number of arguments of template $s$
        \State $A \gets \generategroups(s, \blocks)$
        \ForAll{$(B_1, \dots, B_n) \in A$}
        \State $A' \gets \findassignment(s, (B_1,\dots,B_n))$
        \ForAll{$(B_1',\dots,B_n') \in A'$}
            \State $C \gets C \cup \{ c_s(B_1', \dots, B_n') \}$
          \EndFor
        \EndFor
      \EndFor
      \State \Return $C$
    \EndProcedure
\end{algorithmic}
\caption{Learn tabular constraints}
\label{algo:tcl}
\end{algorithm}

\subsubsection{Generating superblock assignments (Step 3a)}
\label{sec:algo:super}
\
% Let there be $b$ blocks and let~$s$ have~$n$ arguments, then a simple generate-and-test method would have to test $b^n$ combinations (given that blocks can be repeated), which can grow very rapidly with $b$, depending on $n$.

@@

Hence, a simple generate-and-test will usually suffice.
%However, the actual implementation is specified for every constraint template separately, allowing different methods to be used based on the type of template.
%For this reason, the \sname system uses generate-and-test to implement $\findassignment$ for most constraint templates.
\luc{test = definition, why use another term ?}
Given a superblock assignment, all disjoint subassignments are generated taking into account the required size, columns or rows.
For every (disjoint) subassignment the exact signature (e.g. subtypes will not have been checked yet) and definition will be tested and all satisfying subassignments are returned.

\begin{algorithm}[tbh]
  \begin{algorithmic}
    \footnotesize
    \Procedure{Subassignments}{$s$, $(B_1, \dots, B_n)$}
    \State $n \gets$ number of arguments of template $s$
    \State $A_{sub} \gets \emptyset$
    \ForAll{$(B_1', \dots, B_n') \textbf{ where } \forall i: B_i' \sqsubseteq B_i  \land \textit{disjoint}_{j \neq i}(B_i', B_j')$}
      \If{$\mathit{Sig}_s(B_1',\dots,B_n') \land \mathit{Def}_s(B_1',\dots,B_n')$}
        \State $A_{sub} \gets A_{sub} \cup \{ (B_1', \dots, B_n') \}$
      \EndIf
    \EndFor
\Return $A_{sub}$
\EndProcedure
\end{algorithmic}
\caption{Generate-and-test for $\findassignment$}
\label{algo:gen-test-subblocks}
\end{algorithm}
